% ══════════════════════════════════════════════════════════════════════════
\section{Five-Dimensional Minimal Polytopes}
\label{sec:minimal-5d}

This section isolates the combinatorial input from
\citet{KreuzerSkarke1997} that controls which combined weight-system
patterns can occur in dimension~5.  We work with the notation of
their Lemma~1 and Corollary~1: a minimal polytope $M$ is described by an
ordered family of good simplices $S_1,\dots,S_\lambda$ such that the
partial unions span minimal polytopes $M_1,\dots,M_\lambda=M$.

\begin{definition}
Fix a Corollary~1 ordering $S_1,\dots,S_\lambda$ of a $5$-dimensional
minimal polytope.  Let $m_\mu$ be the number of vertices of $S_\mu$ that
do not occur in $S_1 \cup \cdots \cup S_{\mu-1}$.  We call
$(m_1,\dots,m_\lambda)$ the \emph{ordered new-vertex profile} of the
chosen decomposition.
\end{definition}

\begin{proposition}[Five-dimensional profile constraint]
\label{prop:5d-profiles}
Let $M$ be a $5$-dimensional minimal polytope and let
$(m_1,\dots,m_\lambda)$ be the ordered new-vertex profile of a Corollary~1
decomposition.  Then each $m_\mu \ge 2$ and
\[
  \sum_{\mu=1}^{\lambda} (m_\mu - 1) = 5.
\]
Consequently, after sorting the profile in weakly decreasing order one
obtains one of the seven possibilities
\[
  [6],\ [5,2],\ [4,3],\ [4,2,2],\ [3,3,2],\ [3,2,2,2],\ [2,2,2,2,2].
\]
The case $[6]$ is the simplex case, hence the single-weight-system case;
all other profiles correspond to genuine combined weight-system patterns.
\end{proposition}

\begin{proof}
Write $k_\mu$ for the number of vertices in $M_\mu$.  By
Corollary~1(a) of \citet{KreuzerSkarke1997},
$\dim M_\mu = k_\mu - \mu$.  Therefore the dimension jump at step $\mu$ is
\[
  \dim M_\mu - \dim M_{\mu-1}
  = (k_\mu - \mu) - (k_{\mu-1} - (\mu-1))
  = (k_\mu - k_{\mu-1}) - 1
  = m_\mu - 1.
\]
Summing these jumps from $M_1$ to $M_\lambda = M$ yields
\[
  5 = \dim M = \sum_{\mu=1}^{\lambda} (m_\mu - 1).
\]
Each $m_\mu$ is at least $2$, because every new simplex contributes at
least one new dimension and hence at least two vertices.  Subtracting $1$
from each entry therefore produces a partition of $5$ into positive
integers.  The only partitions of $5$ are
$5$, $4+1$, $3+2$, $3+1+1$, $2+2+1$, $2+1+1+1$, and $1+1+1+1+1$.
Adding $1$ back to each part gives exactly the seven profiles listed.
\end{proof}

\begin{lemma}[Lower-dimensional input data]
\label{lem:lower-dimensional-input}
Up to relabeling of vertices, the complete $3$-dimensional minimal-polytope
structures are
\[
  \{\{1,2,3,4\}\},\qquad
  \{\{1,2,3\},\{4,5\}\},\qquad
  \{\{1,2,3\},\{1,4,5\}\},\qquad
  \{\{1,2\},\{3,4\},\{5,6\}\},
\]
and the complete $4$-dimensional minimal-polytope structures are
\[
  \{\{1,2,3,4,5\}\},\qquad
  \{\{1,2,3,4\},\{1,2,5,6\}\},\qquad
  \{\{1,2,3,4\},\{1,5,6\}\},
\]
\[
  \{\{1,2,3,4\},\{5,6\}\},\qquad
  \{\{1,2,3\},\{4,5,6\}\},\qquad
  \{\{1,2,3\},\{4,5\},\{6,7\}\},
\]
\[
  \{\{1,2,3\},\{1,4,5\},\{6,7\}\},\qquad
  \{\{1,2,3\},\{1,4,5\},\{1,6,7\}\},\qquad
  \{\{1,2\},\{3,4\},\{5,6\},\{7,8\}\}.
\]
\end{lemma}

\begin{proof}
This is exactly the explicit $n=3$ and $n=4$ classification stated in
Section~4 of \citet{KreuzerSkarke1997}.  We have only rewritten the same
structures in the present brace notation for good simplices.
\end{proof}

\begin{lemma}[Dimension-$5$ extension rule]
\label{lem:5d-extension-rule}
Let $M$ be a $5$-dimensional minimal polytope.
\begin{enumerate}
  \item If Lemma~1 of \citet{KreuzerSkarke1997} selects $n'=4$, then
  $M$ is obtained from one of the $4$-dimensional base structures of
  \cref{lem:lower-dimensional-input} by adjoining a new good simplex
  \[
    S = R \cup \{a,b\},
  \]
  where $a,b$ are new vertices and $R$ is a subset of old vertices with
  $|R| \le 3$.
  \item If Lemma~1 selects $n'=3$, then $M$ is obtained from one of the
  $3$-dimensional base structures of
  \cref{lem:lower-dimensional-input} by adjoining a new good simplex
  \[
    S = R \cup \{a,b,c\},
  \]
  where $a,b,c$ are new vertices and $R$ is a subset of old vertices with
  $|R| \le 1$.
\end{enumerate}
Conversely, every candidate produced in this way satisfies the dimensional
constraint from Lemma~1.  It survives to a valid structure if and only if
it is not excluded by Lemma~2, i.e. if no good simplex has exactly one
vertex that belongs to no other simplex in the spanning collection.
\end{lemma}

\begin{proof}
Lemma~1 says that a non-simplex $5$-dimensional minimal polytope contains
an $n'$-dimensional minimal polytope $M'$ and one further good simplex $S$
such that
\[
  k-k' = 5-n'+1,
  \qquad
  \dim S \le n'.
\]
Since $n' \ge 5/2$, only $n'=4$ and $n'=3$ can occur.

If $n'=4$, then $k-k' = 2$, so the new simplex contributes exactly two new
vertices, say $a,b$.  Because $\dim S \le 4$, the simplex has at most five
vertices, hence at most three old ones.  Therefore
$S = R \cup \{a,b\}$ with $|R| \le 3$.

If $n'=3$, then $k-k' = 3$, so the new simplex contributes exactly three
new vertices, say $a,b,c$.  Because $\dim S \le 3$, the simplex has at
most four vertices, hence at most one old one.  Therefore
$S = R \cup \{a,b,c\}$ with $|R| \le 1$.

The converse statement about the dimensional constraint is immediate from
the same bounds.  The final sentence is exactly Lemma~2: a spanning set of
good simplices is forbidden precisely when one of the simplices has a
unique vertex not shared by any other simplex.
\end{proof}

\begin{theorem}[Complete $5$D structure list]
\label{thm:5d-structures}
Up to relabeling of the vertices, there are exactly $47$ overlap
structures for $5$-dimensional minimal polytopes.  They break up by total
vertex count as follows:
\begin{center}
\begin{tabular}{@{}cr@{}}
\toprule
Vertices $k$ & Number of structures \\
\midrule
6 & 1 \\
7 & 6 \\
8 & 25 \\
9 & 13 \\
10 & 2 \\
\bottomrule
\end{tabular}
\end{center}
Grouping the same $47$ structures by their sorted new-vertex profiles
produces the multiplicities
\[
  [6] : 1,\quad [5,2] : 2,\quad [4,3] : 4,\quad [4,2,2] : 7,
\]
\[
  [3,3,2] : 18,\quad [3,2,2,2] : 13,\quad [2,2,2,2,2] : 2.
\]
Appendix~\ref{app:5d-structures} lists one canonical Corollary~1 ordering
for each structure, together with its overlap matrix
$O_{ij} = |S_i \cap S_j|$.
\end{theorem}

\begin{proof}
By \cref{lem:lower-dimensional-input}, there are exactly four possible
$3$-dimensional base structures and nine possible $4$-dimensional base
structures.  By \cref{lem:5d-extension-rule}, every non-simplex
$5$-dimensional minimal polytope arises from one of these thirteen bases by
one of the following finite operations:
\begin{enumerate}
  \item choose a $4$D base and adjoin a simplex $R \cup \{a,b\}$ with
  $|R| \le 3$;
  \item choose a $3$D base and adjoin a simplex $R \cup \{a,b,c\}$ with
  $|R| \le 1$.
\end{enumerate}
The simplex case itself contributes the unique structure with profile
$[6]$.

Thus the problem has been reduced to a completely finite search.  For each
base structure, one runs through all allowed subsets $R$ of old vertices,
forms the corresponding candidate spanning collection of good simplices,
and discards the candidate exactly when Lemma~2 is violated.  The
surviving candidates are then quotient by the obvious notion of
isomorphism: permutation of the simplices together with relabeling of the
vertices.  Equivalently, one keeps one canonical representative of each
overlap hypergraph.

This search is complete because every $5$D minimal polytope must arise from
Lemma~1 and therefore from one of the two cases above.  It is correct
because the only combinatorial obstructions used by Kreuzer--Skarke at this
stage are precisely the dimensional bound from Lemma~1 and the uniqueness
obstruction from Lemma~2.  No surviving candidate is identified with a
different one unless their overlap hypergraphs are isomorphic, since the
canonicalization step quotients exactly by vertex relabeling and simplex
permutation.

Carrying out that finite search yields exactly the $47$ canonical overlap
structures listed in Appendix~\ref{app:5d-structures}.  Counting them by
total number of vertices gives
\[
  1,\ 6,\ 25,\ 13,\ 2
\]
for $k=6,7,8,9,10$ respectively, and counting them by sorted new-vertex
profile gives
\[
  [6] : 1,\quad [5,2] : 2,\quad [4,3] : 4,\quad [4,2,2] : 7,
\]
\[
  [3,3,2] : 18,\quad [3,2,2,2] : 13,\quad [2,2,2,2,2] : 2.
\]
The finite search itself, including these exact counts, is machine-checked
in \texttt{lean/MinimalPolytopes5D/Structures.lean} by the theorems
\texttt{canonicalFiveDimensionalStructures\_length},
\texttt{countsByVertex\_correct},
\texttt{countsByProfile\_correct}, and
\texttt{sortedProfile\_allowed}.
\end{proof}

\begin{remark}
The appendix records one admissible Corollary~1 ordering for each overlap
hypergraph.  Some hypergraphs admit more than one such ordering and hence
more than one ordered new-vertex profile.  The profile in the appendix is
therefore a canonical choice, not an invariant attached to the hypergraph
itself.  The arithmetic core of \cref{prop:5d-profiles} is formalized in
Lean~4 in the file \texttt{lean/MinimalPolytopes5D/Profiles.lean}, and the
finite recursive search from the explicit $3$D and $4$D base lists to the
full set of $47$ canonical $5$D structures is formalized in
\texttt{lean/MinimalPolytopes5D/Structures.lean}.
\end{remark}

\subsection{Heuristic candidate counts by geometry type}
\label{subsec:5d-cws-estimates}

The overlap classification also gives a simple first-order model for how
the dimension-$5$ search should distribute across the $47$ geometry types.
For a geometry type $T$, let
\[
  \sigma(T) = \{\, |S_1|, \dots, |S_\lambda| \,\}
\]
be the multiset of simplex sizes read off from the canonical Corollary~1
ordering in Appendix~\ref{app:5d-structures}.  Let $m_k(T)$ be the
multiplicity of $k$ in $\sigma(T)$.  We anchor the estimate by the known
single-weight-system pool sizes
\begin{equation}
\label{eq:arity-pool-sizes}
  A_2 = 1,\qquad A_3 = 3,\qquad A_4 = 95,\qquad
  A_5 = 184{,}026,\qquad A_6 = 183{,}000{,}000{,}000.
\end{equation}
Here $A_2$ is the unique $2$-weight system $(1,1)$, $A_3$ is the number of
single weight systems in dimension $2$, $A_4$ is the number of single
weight systems in dimension $3$, $A_5$ is the number of single weight
systems in dimension $4$, and $A_6$ is the recovered size of the
$6$-weight pool in dimension $5$.  The published lower-dimensional values
$A_3$, $A_4$, and $A_5$ are listed on the Kreuzer--Skarke classification
page~\cite{KreuzerSkarkeCYcsws}.

If one ignores the detailed overlap matrix and only asks how many multisets
of constituent single weight systems have the required arities, then the
natural estimate is the product of multiset coefficients
\begin{equation}
\label{eq:geometry-type-estimate}
  \widehat N(T)
  = \prod_{k=2}^{6} \binom{A_k + m_k(T) - 1}{m_k(T)}.
\end{equation}
This is a multiset count rather than an ordinary product because a geometry
type may use the same single weight system more than once.  For example,
types $12$ and $13$ use simplex sizes $(5,4,4)$, so
\[
  \widehat N(12) = \widehat N(13)
  = A_5 \binom{A_4 + 1}{2}
  = 184{,}026 \binom{96}{2}
  = 839{,}158{,}560.
\]
By construction, \cref{eq:geometry-type-estimate} depends only on the
simplex-size multiset $\sigma(T)$, not on the overlap matrix itself; hence
distinct geometry types with the same simplex-size multiset receive the
same estimate.

The same baseline calculation can be calibrated against the published
dimension-$4$ counts~\cite{KreuzerSkarkeCYcsws}.  For the four families
$4{+}4$, $3{+}4$, $3{+}3$, and $3{+}3{+}3$ one obtains
\[
  \binom{95+1}{2} = 4{,}560,\qquad 3 \cdot 95 = 285,\qquad
  \binom{3+1}{2} = 6,\qquad \binom{3+2}{3} = 10,
\]
whereas the published exact counts are $16{,}040$, $1{,}122$, $6$, and
$36$.  Thus the ratios exact/baseline are $3.52$, $3.94$, $1.00$, and
$3.60$, respectively.  We therefore interpret $\widehat N(T)$ as a
workload estimate anchored to the available single-weight pools, not as an
exact theorem about the final emitted CWS counts.

Using \cref{eq:arity-pool-sizes,eq:geometry-type-estimate}, and fixing the
simplex case $T=1$ to the recovered value $A_6 = 183 \times 10^9$, one
obtains the estimates in \cref{tab:5d-cws-estimates}.  Under this model,
the largest genuine combined type is $T=3$, with estimate
$16{,}932{,}876{,}351$, followed by types $12$ and $13$ at
$839{,}158{,}560$ each.

\begingroup
\small
\setlength{\tabcolsep}{6pt}
\renewcommand{\arraystretch}{1.05}
\begin{longtable}{@{}p{3.8cm}>{\raggedright\arraybackslash}p{5.7cm}r@{}}
\caption{Arity-only heuristic estimates for the $5$D geometry types.  Rows are grouped when the numerical value of \cref{eq:geometry-type-estimate} coincides.}
\label{tab:5d-cws-estimates}\\
\toprule
Type IDs & Simplex sizes & $\widehat N(T)$ \\
\midrule
\endfirsthead
\toprule
Type IDs & Simplex sizes & $\widehat N(T)$ \\
\midrule
\endhead
\bottomrule
\endfoot
1 & $(6)$ & 183,000,000,000 \\
3 & $(5,5)$ & 16,932,876,351 \\
12, 13 & $(5,4,4)$ & 839,158,560 \\
20, 25, 26, 27, 29 & $(5,4,3)$ & 52,447,410 \\
6, 15 & $(5,4)$, $(5,4,2)$ & 17,482,470 \\
43 & $(5,3,3,3)$ & 1,840,260 \\
21, 35, 40 & $(5,3,3)$, $(5,3,3,2)$ & 1,104,156 \\
4, 36, 38 & $(5,3)$, $(5,3,2,2)$ & 552,078 \\
2 & $(5,2)$ & 184,026 \\
11, 14 & $(4,4,4)$ & 147,440 \\
24, 28 & $(4,4,3)$ & 13,680 \\
7, 8, 10 & $(4,4)$, $(4,4,2)$ & 4,560 \\
45 & $(4,3,3,3)$ & 950 \\
22, 30, 31, 32, 34, 42 & $(4,3,3)$, $(4,3,3,2)$ & 570 \\
5, 16, 17, 18, 33 & $(4,3)$, $(4,3,2)$, $(4,3,2,2)$ & 285 \\
9, 46 & $(4,2,2)$, $(4,2,2,2,2)$ & 95 \\
44 & $(3,3,3,3)$ & 15 \\
23, 41 & $(3,3,3)$, $(3,3,3,2)$ & 10 \\
19, 39 & $(3,3,2)$, $(3,3,2,2)$ & 6 \\
37 & $(3,2,2,2)$ & 3 \\
47 & $(2,2,2,2,2)$ & 1 \\
\end{longtable}
\endgroup